---
title: "Theming with bslib and thematic"
author: "Nicolas cabello"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

The initial steps involve loading necessary libraries and data
(pml.training and pml.testing). The data is then cleaned by removing
columns with near-zero variance and missing values, ensuring that the
training data is ready for model

```{r}
library(caret)
library(corrplot)
library(tidyverse)
library(randomForest)
library(FNN)
library(cluster)
library(kernlab)
library(rattle)
library(ggplot2) 

```

```{r}
#con varianza cercana a 0  NAS fuera

testing<- read.csv("pml-testing.csv")
training<- read.csv("pml-training.csv")
training <- training[,-c(1:7)] 
training <- training[, colSums(is.na(training)) == 0]
depured0<- nearZeroVar(training)
trainingd0<-training[,-depured0]
```

After preprocessing, the dataset is split into training and validation
sets using stratified sampling. This is done to ensure that each class
is represented proportionally in both sets.

```{r}
#ok now the split
set.seed(1542) 
traiN <- createDataPartition( trainingd0$classe,
p=0.75,list = F)
trainingR<-trainingd0[traiN,]
validationR<-trainingd0[-traiN,]
```

The first model built is a Random Forest classifier. Cross-validation
(CV) is performed with 5 folds using the trainControl function, and the
model is trained using the train function with the 'rf' method. The
number of trees is set to 50.

```{r}
#modeling rf library(caret) rfcv \<-
trainControl(method = "cv", number = 5) 
rfcvmodel <- train(classe ~
.,data =trainingR, method = "rf", trControl = rfcv,tuneLength =
5,ntree=50) 
print(rfcvmodel)

test_predictions <- predict(rfcvmodel, validationR)

cmrfcv <- confusionMatrix(test_predictions,
as.factor(validationR$classe))

cmrfcv 
# cmrfcv Confusion Matrix and Statistics

          Reference

Prediction A B C D E A 1395 5 0 0 0 B 0 943 5 0 1 C 0 1 847 10 1 D 0 0 3
792 0 E 0 0 0 2 899

Overall Statistics

               Accuracy : 0.9943          
                 95% CI : (0.9918, 0.9962)
    No Information Rate : 0.2845          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.9928          
                                          

Mcnemar's Test P-Value : NA

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E

Sensitivity 1.0000 0.9937 0.9906 0.9851 0.9978 Specificity 0.9986 0.9985
0.9970 0.9993 0.9995 Pos Pred Value 0.9964 0.9937 0.9860 0.9962 0.9978
Neg Pred Value 1.0000 0.9985 0.9980 0.9971 0.9995 Prevalence 0.2845
0.1935 0.1743 0.1639 0.1837 Detection Rate 0.2845 0.1923 0.1727 0.1615
0.1833 Detection Prevalence 0.2855 0.1935 0.1752 0.1621 0.1837 Balanced
Accuracy 0.9993 0.9961 0.9938 0.9922 0.9986
```

A Decision Tree classifier is built next. The same CV settings are used
as with Random Forest, and the 'rpart' method is employed.

```{r}

#arbol de decision 
rpartcvmodel <- train(classe~., data=trainingR,method="rpart", trControl = rfcv, tuneLength = 10)
fancyRpartPlot(rpartcvmodel\$finalModel) rpartcvmodel

test_predictions \<- predict(rpartcvmodel, validationR)
cmrpartcv \<-confusionMatrix(test_predictions, as.factor(validationR\$classe))
cmrpartcv #mrpartcv Confusion Matrix and Statistics

          Reference

Prediction A B C D E A 1135 178 16 74 20 B 20 464 60 24 49 C 73 112 634
134 109 D 148 160 122 545 140 E 19 35 23 27 583

Overall Statistics

               Accuracy : 0.6854          
                 95% CI : (0.6722, 0.6983)
    No Information Rate : 0.2845          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.6029          
                                          

Mcnemar's Test P-Value : \< 2.2e-16

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E

Sensitivity 0.8136 0.48894 0.7415 0.6779 0.6471 Specificity 0.9179
0.96131 0.8943 0.8610 0.9740 Pos Pred Value 0.7976 0.75203 0.5970 0.4888
0.8486 Neg Pred Value 0.9253 0.88687 0.9425 0.9316 0.9246 Prevalence
0.2845 0.19352 0.1743 0.1639 0.1837 Detection Rate 0.2314 0.09462 0.1293
0.1111 0.1189 Detection Prevalence 0.2902 0.12582 0.2166 0.2274 0.1401
Balanced Accuracy 0.8658 0.72513 0.8179 0.7694 0.8105
```

A Gradient Boosted Machine model is also trained using the same CV
settings as the previous models.

```{r}

#gradient boosted gbmcvmodel \<-
train(classe\~., data=trainingR, method="gbm", trControl = rfcv,
tuneLength = 4, verbose = F) 
test_predictions \<- predict(gbmcvmodel,
validationR) 
cmcvgbm \<- confusionMatrix(test_predictions,
as.factor(validationR\$classe))

cmcvgbm #cmcvgbm Confusion Matrix and Statistics

          Reference

Prediction A B C D E A 1384 12 0 0 1 B 7 929 18 0 1 C 3 8 825 11 5 D 0 0
11 789 6 E 1 0 1 4 888

Overall Statistics

               Accuracy : 0.9819          
                 95% CI : (0.9777, 0.9854)
    No Information Rate : 0.2845          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.977           
                                          

Mcnemar's Test P-Value : NA

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E

Sensitivity 0.9921 0.9789 0.9649 0.9813 0.9856 Specificity 0.9963 0.9934
0.9933 0.9959 0.9985 Pos Pred Value 0.9907 0.9728 0.9683 0.9789 0.9933
Neg Pred Value 0.9969 0.9949 0.9926 0.9963 0.9968 Prevalence 0.2845
0.1935 0.1743 0.1639 0.1837 Detection Rate 0.2822 0.1894 0.1682 0.1609
0.1811 Detection Prevalence 0.2849 0.1947 0.1737 0.1644 0.1823 Balanced
Accuracy 0.9942 0.9862 0.9791 0.9886 0.9920
```

The models' accuracy values are compared, and the Random Forest model is
found to be the most accurate among the three.

```{r}

# Print the acuracy
cat("Accuracy:",
cmrfcv$overall['Accuracy'], "RF", cmrpartcv$overall['Accuracy'],"rpart",
cmcvgbm\$overall['Accuracy'], "gbm")
#Accuracy: 0.9942904 RF 0.6853589 rpart 0.9818515 gbm\> 
#random forest more accurate
```

testeando el set

```{r}

predictionsforproject\<- predict(rfcvmodel,testing)
predictionsforproject plot(rfcvmodel) 
\# results B A B A A E D B A A B C B A E E A B B B
```
